---
title: "esm206_lab5"
author: "Caroline Shepherd"
date: "10/26/2021"
output: 
  html_document:
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Confidence Intervals

In this section, we'll learn to use the 't.test()' function as a way to return the confidence interval using the t-distribution.

## Example 1

Create a mock sample of observation values, and find the confidence interval. Then we'll learn how to report that.

```{r}
pika_mass <- c(82, 94, 110, 70, 86, 99, 102, 366)

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_confidence_interval <- t.test(pika_mass)

```

### Below is a risky and tedious way to write out your data

Mean pika mass at Paiute Pass is 91.9g $\pm$ 13.5g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g.

### Below is a high initial investment, but a safer and better way

In-line code referencing to stored obejcts!

Use a single backtick on either side of a lowercase r to create an in-line bit of code.
(backtick is under esc key)

Mean pika mass at Paiute Pass is `r mean(pika_mass)`

or `r pika_mean` because we already calculated and renamed mean pika mass

for rounding, use - 

Mean Pika mass at Paiute Pass is`r round(pika_mean, 1)` g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`). With a 95% confidence interval of [`r pika_confidence_interval$conf.int[1]`, `r pika_confidence_interval$conf.int[2]`] g.

- In console, type 'pika_confidence_interval$p.value' to find p value
  Type pika_confidence_interval$estimate
  pika_confidence_interval$conf_int
  
  
# Two-sample t-test
  
Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference = 0)

- H0: Means difference = 0
- HA: Means difference is NOT = 0

This is a two-sided t-test becuase no directionality is implied.

## Example 1

```{r}
oak_trees <- c(29, 19, 22, 30 ,35, 16, 20, 7)

pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)

```

Is there a significant difference in mean heights for oak and pine trees?

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

The p-value of `r trees_t$p.value` means that the mean of tree height between pine trees and oak trees is significantly different and we can reject the null hypothesis.

The value is 0.018, so there is a 1.8% chance of finding sample means at least as different as those I found if they were drawn from populations with the same mean height.

## Example 2

using the 'mpg' dataset  to compare city gas mileage for SUVs and compact cars.

use view(mpg) in console to view dataset

use unique(mpg$class) to see unique classes in the dataset

```{r}
# create a new subset called compact_suv that only contains observations where the class is suv OR compact

compact_suv <- mpg %>% 
  filter(class %in% c("suv", "compact"))
```

```{r}
ggplot(data = compact_suv, aes(x = cty)) +
  geom_histogram() +
  facet_wrap(~ class)

ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq() +
  facet_wrap(~ class)
```







